Results_paper1

### 2019-05-01 Paper1_results.Rmd
updating
Updated

## Wavelets
Wavelet anlyses showing the temperature variation between sites along the west coast for the years 2010 - 2014. Working with remotely-sensed SST and *in-situ* collected coastal seawater temperature, in the analyses bellow I aim to show how the intensity and duration of upwelling events vary between sites along the west coast of South Africa. It is important to remember that each of the datasets used within this study obtained SSTs at different resolutions.

@@

# Wavelet analyses

```{r}
# When doing a wavelet analyses on the SACTN dataset the "insitu_temp" column needed to be renamed to "temp" for the rest of the code to run. The rest of this code is kept constant for each of the datasets (Remotely-sensed SST and In-situ seawater temperature)

@@

The aim of this is to find the temperature collected at different distances from the coast (distances = 15km, 25km, 35km and 50km).

@@

transect.pixel <- function(site, distances){
  # Extract coordinates
  coords <- data.frame(lon = site$lon, lat = site$lat)
  # Find lon/ lats every X metres 
  pixels <- data.frame()
  # deep <- 999
  # distance_multiplier <- 1
  # while(deep > isobath){
  for(i in 1:length(distances)){
    # Add index to show distance from coast
    # Find depth at distance step
    coords2 <- as.data.frame(destPoint(p = coords, b = site$heading, d = distances[i]))
    sitesIdx <- knnx.index(sa_bathy[,1:2], as.matrix(coords2), k = 1)
    bathy2 <- sa_bathy[sitesIdx,]
    bathy2 <- bathy2[complete.cases(bathy2$depth),]
    bathy3 <- data.frame(site = site$site, lon = bathy2$lon, lat = bathy2$lat, 
                         heading = site$heading, 
                         # depth = bathy2$depth,
                         distance = distances[i])
    pixels <- rbind(pixels, bathy3)
    coords <- coords2
    # distance_multiplier <- distance_multiplier +1 
    # deep <- bathy2$depth
  }
  # pixels2 <- pixels[pixels$depth > -999,] 
  if(nrow(pixels) < 1){
    pixels <- data.frame(site, depth = NA)
  }else{
    pixels <- pixels
  }
  return(pixels)
}

# Pixel points
site_pixels <- data.frame()
for(i in 1:length(west$site)){
  site <- site_transects[i,]
  site_pixel <- transect.pixel(site, c(10000, 20000, 30000, 40000, 50000)) # RWS: The desired distances
  site_pixels <- rbind(site_pixels, site_pixel)
}

@@

## Create a visualisation of the sites at the different distances from the coastline

```{r}
load("Data_P1/south_africa_coast.RData")
load("Data_P1/bathy/sa_bathy.RData")
load("Data_P1/bathy/bathy.RData")
load("Data_P1/site_list_sub.Rdata")
load("Data_P1/site_pixels.RData")
load("Data_P1/MUR.RData")
names(south_africa_coast)[1] <- "lon"

@@

# Define plotting parameters
sa_lats <- c(-37, -27); sa_lons <- c(14, 50)

site_map <- ggplot(data = south_africa_coast, aes(x = lon, y = lat)) + bw_update +
  geom_raster(data = MUR, aes(x = lon, y = lat, fill = temp)) +
   geom_polygon(data = south_africa_coast, aes(x = lon, y = lat, group = group),
                fill = "grey", colour = "grey", size = 0.1, show.legend = FALSE) +
  geom_point(data = site_list_sub, aes(x = lon, y = lat), alpha = 0.8, size = 3) +
  geom_point(data = site_pixels, aes(x = lon, y = lat), colour = "white", shape = 0, alpha = 0.8, size = 2.1) +
  # geom_rect(data = bbox, aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax), 
  #           alpha = 0.4, colour = "red", size = 0.1, linetype = 1) +
  # scale_x_continuous(limits = sa_lons, expand = c(0, 0), breaks = seq(15, 35, 5)) +
  # scale_y_continuous(limits = sa_lats, expand = c(0, 0), breaks = seq(-35, -30, 5)) +
  scale_fill_viridis(
    option = "magma",
    name = "Temperature (°C)",
    alpha = 0.8) +
  #scale_fill_viridis(limits = limits, expression(paste("Temperature (°C)"))) +
  xlab("") + ylab("") +
  annotate("text", label = "INDIAN\nOCEAN", x = 32.00, y = -35.0,
             size = 5, angle = 0, colour = "black") +
    annotate("text", label = "ATLANTIC\nOCEAN", x = 17.00, y = -35.0,
             size = 5, angle = 0, colour = "black") +
    annotate("text", label = "Benguela", x = 16.0, y = -31.7,
             size = 6, angle = 302, colour = "black") +
    annotate("text", label = "Agulhas", x = 31.7, y = -31.7,
             size = 6, angle = 50, colour = "black") +
  guides(fill = guide_colourbar()) +
  scale_x_continuous(expand = c(0, 0),
                       labels = scales::unit_format(unit = "°E", sep = "")) +
    scale_y_continuous(expand = c(0, 0),
                       labels = scales::unit_format(unit = "°S", sep = "")) +
  theme(panel.background = element_rect(fill = "ivory", colour = NA),
        panel.border = element_rect(colour = "black", size = 0.5),
        panel.grid.minor = element_line(colour = "NA"),
        panel.grid.major = element_line(colour = "ivory", size = 0.2, linetype = "dotted"),
        legend.text = element_text(size = 8),
        legend.title = element_text(size = 8),
        legend.key = element_blank(),
        legend.background = element_blank(),
        axis.ticks = element_line(size = 0.5))
site_map

@@

# Plotting the wind data - Waverose diagrams

@@

              paddle=FALSE, type="station", width=2))
)
png("Figures/wind.png", width = 8, height = 5, pointsize = 6)
?windRose
# insitu_wind_2 <- insitu_wind %>% 
#   filter(site == "Port Nolloth")
# 
# insitu_wind_3 <- insitu_wind %>% 
#   filter(site == "Lamberts Bay")
# 
# plotPN <- plot(as.factor(format(insitu_wind_2$date, "%Y-%m")), insitu_wind_2$insitu_temp, col = "lightpink")
# plotLB <- plot(as.factor(format(insitu_wind_3$date, "%Y-%m")), insitu_wind_3$insitu_temp, col = "lightpink")
#        
# ?plot
```

@@

Here I want to test using generalised linear hypotheses if there is a difference in the intensity of upwellling events between the insitu and satellite data.
Do the in situ and satelllite derived data yield the same number of events.

library(multcomp) # for generalized linear hypothesis test (glht)

@@

load("Data_P1/SACTN_fyears.RData")
load("Data_P1/MUR_fyears.RData")
load("Data_P1/OISST_fyears.RData")
load("Data_P1/G1SST_fyears.RData")
load("Data_P1/K10_fyears.RData")

@@

R2 <- data.frame()
for(i in 1:length(levels(SACTN_fyears$site))) {
  x <- subset(SACTN_fyears, site == levels(SACTN_fyears$site)[i])
  y <- subset(OISST_fyears, site == levels(OISST_fyears$site)[i])
  x <- x[x$date %in% y$date,]
  y <- y[y$date %in% x$date,]
  z <- data.frame(site = as.character(x$site[1]), R2 = coef(lm(y$temp~x$temp))[2],na.action=na.exclude)
  R2 <- rbind(R2, z)
}
```

## EOF's

@@

Mond:
- Try PCA
- map it/plot
- dave upwelling index

Tuesday
- Do exercises

>>>>

## Wavelets
Wavelet anlyses showing the temperature variation between sites along the west coast for the years 2010 - 2014. Working with remotely-sensed SST and *in-situ* collected coastal seawater temperature, in the analyses below I aim to show how the intensity and duration of upwelling events vary between sites along the west coast of South Africa. It is important to remember that each of the datasets used within this study obtained SSTs at different resolutions.

@@

# Wavelet analyses.
    # This creates a visualisation looking at the intensity and duration of upwelling. 
    # Here we test how the intensity and duration of an upwelling event vary between sites along the west coast of SA
    # With the hypotheses that sites closer to the upwelling centre will show a more intense, longer upwelling event

```{r}
# When doing a wavelet analyses on the SACTN dataset the "insitu_temp" column needed to be renamed to "temp" for the rest of the code to run. 
# The rest of this code is kept constant for each of the datasets (Remotely-sensed SST and In-situ seawater temperature)

@@

The aim of this is to find the temperature collected at different distances from the coast (distances = 10km, 20km, 30km, 40km and 50km).

@@

transect.pixel <- function(site, distances){
  # Extract coordinates
  coords <- data.frame(lon = site$lon, lat = site$lat)
  # Find lon/ lats every X metres 
  pixels <- data.frame()
  # deep <- 999
  # distance_multiplier <- 1
  # while(deep > isobath){
  for(i in 1:length(distances)){
    coords2 <- as.data.frame(destPoint(p = coords, b = site$heading, d = distances[i]))
    sitesIdx <- knnx.index(sa_bathy[,1:2], as.matrix(coords2), k = 1)
    bathy2 <- sa_bathy[sitesIdx,]
    bathy2 <- bathy2[complete.cases(bathy2$depth),]
    bathy3 <- data.frame(site = site$site, lon = bathy2$lon, lat = bathy2$lat, 
                         heading = site$heading, 
                         distance = distances[i])
    pixels <- rbind(pixels, bathy3)
    coords <- coords2
  }
  if(nrow(pixels) < 1){
    pixels <- data.frame(site, depth = NA)
  }else{
    pixels <- pixels
  }
  return(pixels)
}

# Pixel points
site_pixels <- data.frame()
for(i in 1:length(west$site)){
  site <- site_transects[i,]
  site_pixel <- transect.pixel(site, c(10000, 20000, 30000, 40000, 50000)) # RWS: fixed error
  site_pixels <- rbind(site_pixels, site_pixel)
}

@@

## Create a visualisation of the sites at the different distances from the coastline. Mapping the study area

```{r}
load("Data_P1/south_africa_coast.RData")
# load("Data_P1/bathy/sa_bathy.RData")
# load("Data_P1/bathy/bathy.RData")
load("Data_P1/site_list_sub.Rdata")
load("Data_P1/site_pixels.RData")
load("Data_P1/MUR.RData")
names(south_africa_coast)[1] <- "lon"

@@

# Define plotting parameters
sa_lats <- c(-37, -27); sa_lons <- c(14, 25)

site_map <- ggplot(data = south_africa_coast, aes(x = lon, y = lat)) + bw_update +
  geom_raster(data = MUR, aes(x = lon, y = lat, fill = temp)) +
   geom_polygon(data = south_africa_coast, aes(x = lon, y = lat, group = group),
                fill = "grey", colour = "grey", size = 0.1, show.legend = FALSE) +
  geom_point(data = site_list_sub, aes(x = lon, y = lat), alpha = 0.8, size = 3) +
  geom_point(data = site_pixels, aes(x = lon, y = lat), colour = "white", shape = 0, alpha = 0.8, size = 2.1) +
  coord_equal(xlim = c(15, 25), ylim = c(-36,-26), expand = 0) +
  # geom_rect(data = bbox, aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax),
  #           alpha = 0.4, colour = "red", size = 0.1, linetype = 1) +
  # scale_x_continuous(limits = sa_lons, expand = c(0, 0), breaks = seq(15, 35, 5)) +
  # scale_y_continuous(limits = sa_lats, expand = c(0, 0), breaks = seq(-35, -30, 5)) +
  scale_fill_viridis(
    option = "magma",
    name = "Temperature (°C)",
    alpha = 0.8) +
  xlab("") + ylab("") +
  # annotate("text", label = "INDIAN\nOCEAN", x = 32.00, y = -35.0,
  #            size = 4, angle = 0, colour = "black") +
    annotate("text", label = "ATLANTIC\nOCEAN", x = 17.00, y = -35.0,
             size =4, angle = 0, colour = "black") +
    annotate("text", label = "Benguela", x = 16.0, y = -31.7,
             size = 5, angle = 302, colour = "black") +
    # annotate("text", label = "Agulhas", x = 31.7, y = -31.7,
    #          size = 5, angle = 50, colour = "black") +
  guides(fill = guide_colourbar()) +
  scale_x_continuous(expand = c(0, 0),
                       labels = scales::unit_format(unit = "°E", sep = "")) +
    scale_y_continuous(expand = c(0, 0),
                       labels = scales::unit_format(unit = "°S", sep = "")) +
  theme(panel.background = element_rect(fill = "ivory", colour = NA),
        panel.border = element_rect(colour = "black", size = 0.5),
        panel.grid.minor = element_line(colour = "NA"),
        panel.grid.major = element_line(colour = "ivory", size = 0.2, linetype = "dotted"),
        legend.text = element_text(size = 8),
        legend.title = element_text(size = 8),
        legend.key = element_blank(),
        legend.background = element_blank(),
        axis.ticks = element_line(size = 0.5))
site_map

@@

# Plotting the wind data - windrose diagrams
  # Here I produce a wind rose diagram showing wind action or wind patterns during the upwelling period

@@

              paddle=FALSE, type="station", width=2))
)
# png("Figures/wind.png", width = 8, height = 5, pointsize = 6)
?windRose

# insitu_wind_2 <- insitu_wind %>% 
#   filter(site == "Port Nolloth")
# 
# insitu_wind_3 <- insitu_wind %>% 
#   filter(site == "Lamberts Bay")
# 
# plotPN <- plot(as.factor(format(insitu_wind_2$date, "%Y-%m")), insitu_wind_2$insitu_temp, col = "lightpink")
# plotLB <- plot(as.factor(format(insitu_wind_3$date, "%Y-%m")), insitu_wind_3$insitu_temp, col = "lightpink")
#        
# ?plot
```

@@

First I calculate the distance between the sites along the wc using the deg2rad function which transforms degrees to radians: Transforms between angles in degrees and radians. Thereafter I run a three way ANOVA. This was primarily done to test whether a significant difference in temperature exist between sites on a seasonal, yearly bases. Here the distance between the sites could also play a role. But distance is mainly used when compaing the intensity and duration of an upwelling event given the distance between the sites and its location (location: at an upwelling centre or further away)

```{r}
load("Data_P1/sa_coast.Rdata")
load("Data_P1/site_list_sub.Rdata")
load("Data_P1/SACTN_fyears.RData")
load("Data_P1/MUR_fyears.RData")
load("Data_P1/OISST_fyears.RData")
load("Data_P1/G1SST_fyears.RData")
load("Data_P1/K10_fyears.RData")
SACTN_fyears <- SACTN_fyears %>% 
  dplyr::rename(temp = insitu_temp)

options(scipen=999) # Forces R to not use exponential notation
sa_coast_prep <- sa_coast %>% 
  mutate(site = 1:nrow(.),
         X = deg2rad(X),
         Y = deg2rad(Y)) %>% 
  select(site, X, Y)

sa_coast_dist <- as.data.frame(round(PairsDists(sa_coast_prep), 2)) %>% 
  mutate(lon = sa_coast$X,
         lat = sa_coast$Y) %>% 
  dplyr::rename(dist = V1) %>%
  select(lon, lat, dist) %>% 
  mutate(cum_dist = cumsum(dist)) %>% 
  mutate(dist = lag(dist),
         cum_dist = lag(cum_dist))

# Calculating the distances between sites.
# This may be used to show how upwelling temperature varies between sites based on its location, i.e. closer 
# to the upwelling centre and further from the upwelling centre

SACTN_site_dis <- function(df) {
  SACTN_dis <- data.frame()
  df2 <- df %>%
    select(-length) %>%
    droplevels()
  for (i in 1:length(levels(df2$index))) {  
    # for (i in 1:10) {
    SACTN_df_1 <- filter(df2, index == levels(index)[i])
       # Find the cumulative distance of site 1 along the coast
    site_dist_1 <- site_list_sub %>% 
      filter(index == as.character(SACTN_df_1$index)[1])
    site_dist_1 <- knnx.index(data = as.matrix(sa_coast_dist[,1:2]),
                              query = as.matrix(site_dist_1[,5:6]), k = 1)
    site_dist_1 <- sa_coast_dist$cum_dist[site_dist_1]

    for (j in 1:length(levels(df2$index))) {
      # for(j in 1:10) {
      if (i == j) {
        next
      }
      if (j < i) {
        next
      }
      SACTN_df_2 <- filter(df2, index == levels(index)[j])

      # Find the cumulative distance of site 2 along the coast
      site_dist_2 <- site_list_sub %>% 
        filter(index == as.character(SACTN_df_2$index)[1])
      site_dist_2 <- knnx.index(data = as.matrix(sa_coast_dist[,1:2]),
                                query = as.matrix(site_dist_2[,5:6]), k = 1)
      site_dist_2 <- sa_coast_dist$cum_dist[site_dist_2]

      # # The distance between the two sites
      site_dist_3 <- round(abs(site_dist_1-site_dist_2), 0)

      SACTN_df_3 <- left_join(SACTN_df_1, SACTN_df_2, by = "date") %>%
        na.trim() %>% 
      select(date, index.x, temp.x, index.y, temp.y) %>%
        dplyr::rename(index_1 = index.x,
          temp_1 = temp.x,
          index_2 = index.y,
          temp_2 = temp.y) %>% 
        mutate(index_dist = site_dist_3)
      SACTN_dis <- rbind(SACTN_dis, SACTN_df_3)
    }
  }
  return(SACTN_dis)
}
SACTN_site_dist <- SACTN_site_dis(df = SACTN_fyears)

# Adding seasonal column

SACTN_monthly <- function(df) {
  SACTN_split_monthly <- df %>%
  mutate(month = month(date, abbr = T, label = T),
         year = year(date)) %>%
  group_by(index, month, year) %>%
    mutate(season = ifelse(month %in% c("Jan", "Feb", "Mar"), "Summer",       
                           ifelse(month %in% c("Apr", "May", "Jun"), "Autumn",
                                ifelse(month %in% c("Jul", "Aug", "Sep"), "Winter",
                                       ifelse(month %in% c("Oct", "Nov", "Dec"), "Spring","Error")))))
  return(SACTN_split_monthly)
}

# SACTN_season <- SACTN_monthly(df = SACTN_fyears)
# MUR_fyears_season <- SACTN_monthly(df = MUR_fyears)
# OISST_fyears_season <- SACTN_monthly(df = OISST_fyears)
# G1SST_fyears_season <- SACTN_monthly(df = G1SST_fyears)
# K10_fyears_season <- SACTN_monthly(df = K10_fyears)

# save(SACTN_season, file = "SACTN_season.RData")
# save(MUR_fyears_season, file = "MUR_fyears_season.RData")
# save(OISST_fyears_season, file = "OISST_fyears_season.RData")
# save(G1SST_fyears_season, file = "G1SST_fyears_season.RData")
# save(K10_fyears_season, file = "K10_fyears_season.RData")

load("Data_P1/SACTN_season.RData")
load("Data_P1/MUR_fyears_season.RData")
load("Data_P1/OISST_fyears_season.RData")
load("Data_P1/G1SST_fyears_season.RData")
load("Data_P1/K10_fyears_season.RData")

# ANOVA analyses
anova_func <- function(df){
  sites_aov <- aov(temp ~ site * year * season, data = df)
return(sites_aov)
}

SACTN_season <- SACTN_season %>% 
  dplyr::rename(temp = insitu_temp)

SACTN_season_func <- anova_func(df = SACTN_season)
summary(SACTN_season_func)
MUR_season_func <- anova_func(df = MUR_fyears_season)
summary(MUR_season_func)
OISST_season_func <- anova_func(df = OISST_fyears_season)
summary(OISST_season_func)
G1SST_season_func <- anova_func(df = G1SST_fyears_season)
summary(G1SST_season_func)
K10_season_func <- anova_func(df = K10_fyears_season)
summary(K10_season_func)

# Shows a significant different in temperatures between sites for each of the years and between seasons

#    ? This is for all temperatures collected. not only temperatures collected during an upwelling event
#    ? So now I need to find a way to restrict temperatures to only the periods where upwelling occured
#    ? Can use the heatwaveR package to identify the count, intensity and duration but for this I require to know what the upwelling threshold is
#     MHW: threshold- 90th percentile, MCS: Threshold - 10th percentile
```

# Here I want to test using generalised linear hypotheses if there is a difference in the intensity of upwellling events between the insitu and       satellite data.
# Do the in situ and satelllite derived data yield the same number of events. First I add a seasonal column to test seasonal variation between 
library(multcomp) # for generalized linear hypothesis test (glht)

@@

load("Data_P1/SACTN_fyears.RData")
load("Data_P1/OISST_fyears.RData")

@@

R2 <- data.frame()
for(i in 1:length(levels(SACTN_fyears$site))) {
  x <- subset(SACTN_fyears, site == levels(SACTN_fyears$site)[i])
  y <- subset(df, site == levels(df$site)[i])
  x <- x[x$date %in% y$date,]
  y <- y[y$date %in% x$date,]
  z <- data.frame(site = as.character(x$site[1]), R2 = coef(lm(y$temp~x$temp))[2])
  R2 <- rbind(R2, z)
}
```

# Upwelling index
## Determining upwelling index from wind data (WindGuru and SAWS)
## Index Equation from Fielding & Davis 1989 paper

```{r}
load("Data_P1/insitu_wind.RData")
load("Data_P1/SACTN_fyears.RData")
load("Data_P1/MUR_fyears.RData")
load("Data_P1/OISST_fyears.RData")
load("Data_P1/G1SST_fyears.RData")
load("Data_P1/K10_fyears.RData")

# UPWELLING INDEX EQUATION  - FIELDING & DAVIS 1989
# Index = Wind_speed*cos(wind_direction) - 160) [*160 refers to the angle (degrees) of the coast]

#data wrangling - TEMPERATURE
SACTN_fyears_UI <- SACTN_fyears %>% 
  dplyr::rename(temp = insitu_temp) %>%
  group_by(site) %>%
  select("date", "temp")

wind_UI <- insitu_wind %>%
          group_by(site) %>%
          mutate(ui.saws = mean_speed * (cos(dir_circ - 160)))

index <- wind_UI %>%
  group_by(site) %>% 
  select("site", "date","ui.saws") %>% 
  left_join(SACTN_fyears_UI, by = c("date", "site")) %>% 
  mutate(ws.condition = ifelse(ui.saws > 0, "upwelling", "downwelling"))

#Plotting the data
#SAWS data
ggplot(index, aes(x= date, y=ui.saws)) +
  geom_area(aes(fill=ws.condition)) +
  facet_wrap(~site) +
  geom_line() +
  geom_hline(yintercept=0)

# Visualisation
```

### EOF's

@@

To do:
- Try PCA
- plot
- dave upwelling index
-
